---
title: "caladapt-r"
format: html
---

```{r}
# Install caladaptr
#devtools::install_github("ucanr-igis/caladaptr")

# Libraries
library(devtools)
library(caladaptr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(sf)
library(rlist)
```

## Request daily maximum temperature for California school districts

```{r}
# Import school bounds
schools_bounds <- st_read("/capstone/casaschools/schools_data/schools_intersect/schools_intersect_bounds.shp") %>% 
  # Elementary school boundaries covers the same boundary areas as high schools
  # Reduce repition include only elementary and unified
  filter(DistrictTy %in% c("Elementary","Unified"))

# Remove duplicates
schools_bounds <- schools_bounds[!duplicated(schools_bounds$DistrictNa),]

schools_bounds <- st_make_valid(schools_bounds)

#plot(schools_bounds$geometry)

# Empty data frame
percentile <- data.frame()

# Loop across all school districts
for (name in schools_bounds$DistrictNa) {
  
  # Iterate through individual rows at a time
  df <- schools_bounds %>% 
    filter(DistrictNa ==  name)
  
  tryCatch({
    # API request
    request <- ca_loc_sf(loc = df, idfld = "CDSCode") %>% 
      ca_cvar(cvar = "pr") %>% 
      # Historic dataset
      ca_livneh(TRUE) %>% 
      ca_period("day")  %>% 
      ca_years(start = 1961, end = 1990)
    
    school_data <- request %>% 
      # Get values from request
      ca_getvals_tbl(quiet = TRUE) %>% 
      mutate(val = as.numeric(val)) %>%
      # Calculate 98th percentile for each district historic data
     summarize(percentile_98 = quantile(val, probs = 0.98))
    
    school_data$DistrictNa <- name
    
    percentile <- rbind(percentile, school_data)
  }, error = function(e) {
    cat("Error occurred for district:", name, "\n")
    # Print school districts API has difficulty with
    cat("Skipping this district.\n")
    
    
  })
}


```

## Working with Errors

```{r}
## By Counties tracts
counties <- ca_aoipreset_geom("counties")

# School districts that are not being read
schools_ers <- c("Del Norte County Unified","Big Sur Unified", 
                 "Cabrillo Unified", "South San Francisco Unified",
                 "Belmont-Redwood Shores Elementary",
                 "San Bruno Park Elementary", "San Mateo-Foster City",
                 "Horicon Elementary") 

# Filter school bounds to just be the error districts
error_bounds <- schools_bounds %>% 
  filter(DistrictNa %in% c(schools_ers))

# Make valid
error_bounds <- st_make_valid(error_bounds)

# Intersect error bounds to be within CalAdapt census counties
counties_in_school <- st_intersection(counties, error_bounds) %>% 
  select(c("CDCode","DistrictNa"))

# Unify all geoms by district name
data_df <- counties_in_school %>%
  st_cast("POLYGON") %>%
  group_by(DistrictNa) %>%
  summarise(geometry = st_union(geom)) %>%
  st_sf()


```



```{r}
# Empty data frame
percentile_2 <- data.frame()

# Loop across all school districts
for (name in data_df$DistrictNa) {
  
  # Iterate through individual rows at a time
  df <- data_df %>% 
    filter(DistrictNa ==  name)
  
  tryCatch({
    # API request
    request <- ca_loc_sf(loc = df, idfld = "DistrictNa") %>% 
      ca_cvar(cvar = "pr") %>% 
      # Historic dataset
      ca_livneh(TRUE) %>% 
      ca_period("day")  %>% 
      ca_years(start = 1961, end = 1990)
    
    data_2 <- request %>% 
      # Get values from request
      ca_getvals_tbl(quiet = TRUE) %>% 
      mutate(val = as.numeric(val)) %>%
      # Calculate 98th percentile for each district historic data
      summarize(percentile_98 = quantile(val, probs = 0.98))
    
    data_2$DistrictNa <- name
    
    percentile_2 <- rbind(percentile_2, data_2)
  }, error = function(e) {
    cat("Error occurred for district:", name, "\n")
    # Print school districts API has difficulty with
    cat("Skipping this district.\n")
    
  })
}


```


## Second Approach

Clipping the school boundaries with CalAdapt pre disposed counties boundaries 
worked for only 3 school districts. 5 school districts are still"fall outside 
the area covered by Cal-Adapt". To work around this random points will be 
generated in each remaining 5 school districts. These points will be used to determine 
the historical maximum temperature.

```{r}
set.seed(123)

# School districts that are not being read
schools_ers <- c("Big Sur Unified", "South San Francisco Unified",
                 "San Bruno Park Elementary", "San Mateo-Foster City",
                 "Horicon Elementary") 

data_df <- data_df %>% 
  filter(DistrictNa %in% schools_ers)

percentile_3 <- data.frame()


for (name in data_df$DistrictNa){
  print(name)
  
  # create new dataframe with one school district
  school <- data_df %>% 
    filter(DistrictNa == name)
  
  # Generate 50 random points for school district
  school_points <- st_sample(school, 50)
  
  school_points <- st_transform(school_points, crs = "EPSG:4326" ) %>% 
  st_coordinates()
  
  # API request
  tryCatch({
  request <- ca_loc_pt(coords = school_points) %>%
  ca_cvar(cvar = "pr") %>%
  # Historic dataset
  ca_livneh(TRUE) %>%
  ca_period("day")  %>%
  ca_years(start = 1961, end = 1990)
  
  data_3 <- request %>% 
      # Get values from request
      ca_getvals_tbl(quiet = TRUE) %>% 
      mutate(val = as.numeric(val)) %>%
      # Calculate 98th percentile for each district historic data
      summarize(percentile_98 = quantile(val, probs = 0.98))
    
    data_3$DistrictNa <- name
    
    percentile_3 <- rbind(percentile_3, data_3)
  
  
  }, error = function(e) {
    cat("Error occurred for district:", name, "\n")
    # Print school districts API has difficulty with
    cat("Skipping this district.\n")
  
})
}

```


```{r}
# # Join 98th percentile dataframes
 percentile_all <- rbind(percentile, percentile_2, percentile_3)
# 
# 
# # 98th percentile threshold for all school districts
threshold <- percentile_all %>% 
  summarise(percentile98_mean = mean(percentile_98))
# 
# 
 write.csv(threshold, "output_data/threshold_precip.csv")
# 
 paste("Average 98th percentile threshold of yearly rainfall totals for all of California's public school districts:",
      round(threshold$percentile98_mean,3),"in")
# 

```


