---
title: "Accessing Extreme Precipitation Days for all Schools from 2034-64"
format: html
author: "CASAschools Team"
---

```{r}
# Libraries
library(dplyr)
library(tidyr)
library(sf)
library(lubridate)
library(caladaptr)
library(units)
```

Extreme precipitation is calculated as the maximum daily precipitation totals over 0.73inches. This overall threshold was calculated in \`precipitation/caladapt_precip_r.qmd\`. Determining the number of extreme heat days follows a similar approach. Still working with Cal-Adapt's Rpackage the first step is to acquire daily maximum precipitation. Data is requested through Cal-Adapt's API request. Due to the size of the data and the instances of crashing the API request data was extracted at 30-year intervals for the different scenarios. This QMD is for accessing RCP8.5 data from 2000 to 2064.

```{r}
# Load school point data
school_points <- st_read("/capstone/casaschools/schools_data/California_Schools_2022-23/California_Schools_2022-23.shp")
```

```{r}
# Filter to schools are active and remove unnecessary columns
school_points <- school_points %>% filter(Status == "Active") %>% 
  select(c(CDSCode))

```

## RCP 8.5

### MID CENTURY 2035 - 2064

```{r}
# Import LOCA grid 
locagrid_sf <- ca_locagrid_geom()

# Check CRS of both data
st_crs(school_points) == st_crs(locagrid_sf)

# Transform school points to LOCA CRS
school_points <- st_transform(school_points, crs = st_crs(locagrid_sf))

# Spatially join points and grid
ca_schools_loca_sf <- school_points %>% st_join(locagrid_sf) 
ca_schools_loca_sf %>% st_drop_geometry()
```

```{r}
# Unique ID values
# we only have to query 1,619 school points
loca_ids_schools <- ca_schools_loca_sf %>% pull(id) %>% unique()


# Create a point layer for the 1,619 LOCA grid cells that contain schools
loca_ctr_sf <- locagrid_sf %>%
  filter(id %in% loca_ids_schools) %>%
  st_centroid()


# API request 
locaschl_et_cap <- ca_loc_sf(loc = loca_ctr_sf, idfld = "id") %>%
  # CanESM2, CNRM-CM5, HadGEM2-ES, MIROC5
  ca_gcm(gcms[1:4]) %>%
  ca_scenario(c("rcp85")) %>%
  ca_cvar("pr") %>%
  ca_period("day") %>%
  # Mid Century 
  ca_years(start = 2035, end = 2064)

# Check if request has no erros
locaschl_et_cap %>% ca_preflight()

head(locaschl_et_cap)
```

```{r}
#Next, we fetch the data. Because this is still a lot of locations, we’ll use ca_getvals_db() to save it in a database:
data_dir_precip_mid_85 <- tools::R_user_dir("pr_mid_85", which = "data")
schools_dir_precip_mid_85 <- file.path(data_dir_precip_mid_85, "schools_pr_mid_85") %>% normalizePath(mustWork = FALSE)
if (!file.exists(schools_dir_precip_mid_85)) dir.create(schools_dir_precip_mid_85, recursive = TRUE)

## Define a new SQLite file name
locaschl_fn_mid85 <- file.path(schools_dir_precip_mid_85, "loca_schl.sqlite") %>% normalizePath(mustWork = FALSE)


## Fetch data
locaschl_et_rtbl <- locaschl_et_cap %>%
  ca_getvals_db(db_fn = locaschl_fn_mid85, db_tbl = "locaschl_et", new_recs_only = TRUE, quiet = TRUE)


head(locaschl_et_rtbl)
```

```{r}
# Transform dataset 
locaschl_et_rtbl_1 <- locaschl_et_rtbl %>% 
  mutate(month = lubridate::month(dt),
         year = lubridate::year(dt),
         val = as.numeric(val))

# Set threshold determined in 'caladapt_precip_r.qmd' and calculate days over
locaschl_et_rtbl_1 <- locaschl_et_rtbl_1 %>%
   mutate(threshold = ifelse(val > 0.000216, 1, 0)) %>%
   # Find total number of days by month and year
   group_by(id, month, year) %>%
   summarise(count = sum(threshold))
# 
# # View results
 head(locaschl_et_rtbl_1)
```

```{r}
# # Find average monthly yearly days above threshold for 4 GCM's
 locaschl_et_rtbl_2 <- locaschl_et_rtbl_1 %>% 
   mutate(average = count/4)

# Rejoin with school points + LOCA id
 ca_schools_et_sf <- ca_schools_loca_sf %>%  
   left_join(locaschl_et_rtbl_2, by = "id", copy = TRUE) 

 head(ca_schools_et_sf)
```

```{r}
# Find yearly extreme precipitation
 ca_schools_et_year <- ca_schools_et_sf %>% 
   group_by(CDSCode, year) %>% 
   summarise(total = sum(count)) %>% 
st_drop_geometry()
#   
 # Find sum total of extreme precipitation days across 1,0008 schools from 2035-2064
 ca_schools <- ca_schools_et_year %>% 
   group_by(CDSCode) %>% 
   summarise(total_35_64 = sum(total)) %>% 
   mutate(total_35_64 = round(total_35_64,digits = 0))
```

```{r}
# Find yearly extreme precipitation
 ca_schools_et_year <- ca_schools_et_sf %>% 
   group_by(CDSCode, year) %>% 
   summarise(total = sum(count)) %>% 
st_drop_geometry()
#   
 # Find sum total of extreme precipitation days across 1,0008 schools from 2035-2064
 ca_schools <- ca_schools_et_year %>% 
   group_by(CDSCode) %>% 
   summarise(total_35_64 = sum(total)) %>% 
   mutate(total_35_64 = round(total_35_64,digits = 0))
```

```{r}
#Export Extreme Precipitation Totals for Mid Century Scenario

# write.csv(ca_schools, "/capstone/casaschools/precipitation/midcentury85.csv")
# 
# write.csv(ca_schools, "precipitation/midcentury85.csv")
# 
# #Export Extreme Precipitation Totals per year for Mid Century Scenario
# 
# write.csv(ca_schools_et_year, "/capstone/casaschools/precipitation/midcentury85_year.csv" )
# 
# write.csv(ca_schools_et_year, "precipitation/midcentury85_year.csv" )
```

### CURRENT CENTURY 2006 - 2034

```{r}
# API request 
 locaschl_et_cap <- ca_loc_sf(loc = loca_ctr_sf, idfld = "id") %>%
#   # CanESM2, CNRM-CM5, HadGEM2-ES, MIROC5
   ca_gcm(gcms[1:4]) %>%
   ca_scenario(c("rcp85")) %>%
   ca_cvar("pr") %>%
   ca_period("day") %>%
#   # Mid Century 
   ca_years(start = 2006, end = 2034)
# 
# # Check if request has no erros
 locaschl_et_cap %>% ca_preflight()
# 
 head(locaschl_et_cap)
```

```{r}
#Next, we fetch the data. Because this is still a lot of locations, we’ll use ca_getvals_db() to save it in a database:
 data_dir_precip_cur_85 <- tools::R_user_dir("pr_cur_85", which = "data")
 schools_dir_precip_cur_85 <- file.path(data_dir_precip_cur_85, "schools_pr_cur_85") %>% normalizePath(mustWork = FALSE)
 if (!file.exists(schools_dir_precip_cur_85)) dir.create(schools_dir_precip_cur_85, recursive = TRUE)
# 
# ## Define a new SQLite file name
 locaschl_fn_cur85 <- file.path(schools_dir_precip_cur_85, "loca_schl.sqlite") %>% normalizePath(mustWork = FALSE)
# 
# 
# ## Fetch data
 locaschl_et_rtbl <- locaschl_et_cap %>%
   ca_getvals_db(db_fn = locaschl_fn_cur85, db_tbl = "locaschl_et", new_recs_only = TRUE, quiet = TRUE)
# 
# 
 head(locaschl_et_rtbl)
```

```{r}
# # Transform dataset 
 locaschl_et_rtbl_1 <- locaschl_et_rtbl %>% 
   mutate(month = lubridate::month(dt),
          year = lubridate::year(dt),
         val = as.numeric(val))
# 
# # Set threshold determined in 'caladapt_precip_r.qmd' and calculate days over
 locaschl_et_rtbl_1 <- locaschl_et_rtbl_1 %>%
    mutate(threshold = ifelse(val > 0.000216, 1, 0)) %>%
#    # Find total number of days by month and year
    group_by(id, month, year) %>%
    summarise(count = sum(threshold))
# # 
# # # View results
  head(locaschl_et_rtbl_1)
```

```{r}
# # Find average monthly yearly days above threshold for 4 GCM's
  locaschl_et_rtbl_2 <- locaschl_et_rtbl_1 %>% 
    mutate(average = count/4)
# 
# # Rejoin with school points + LOCA id
  ca_schools_et_sf <- ca_schools_loca_sf %>%  
   left_join(locaschl_et_rtbl_2, by = "id", copy = TRUE) 
# 
 head(ca_schools_et_sf)
```

```{r}
# Find yearly extreme precipitation
  ca_schools_et_year <- ca_schools_et_sf %>% 
    group_by(CDSCode, year) %>% 
    summarise(total = sum(count)) %>% 
st_drop_geometry()
# #   
#  # Find sum total of extreme precipitation days across 1,0008 schools from 2035-2064
  ca_schools <- ca_schools_et_year %>% 
    group_by(CDSCode) %>% 
    summarise(total_06_34 = sum(total)) %>% 
   mutate(total_06_34 = round(total_06_34,digits = 0))
```

```{r}
#Export Extreme Precipitation Totals for Mid Century Scenario

# write.csv(ca_schools, "/capstone/casaschools/precipitation/curcentury85.csv")
# 
# write.csv(ca_schools, "precipitation/curcentury85.csv")
# 
# #Export Extreme Precipitation Totals per year for Mid Century Scenario
# 
# write.csv(ca_schools_et_year, "/capstone/casaschools/precipitation/curcentury85_year.csv" )
# 
# write.csv(ca_schools_et_year, "precipitation/curcentury85_year.csv" )
```

### HISTORIC / OBSERVED 2000 - 2005

```{r}
# API request 
locaschl_et_cap <- ca_loc_sf(loc = loca_ctr_sf, idfld = "id") %>%
  # CanESM2, CNRM-CM5, HadGEM2-ES, MIROC5
  ca_gcm(gcms[1:4]) %>%
  ca_scenario(c("historical")) %>%
  ca_cvar("pr") %>%
  ca_period("day") %>%
  # Mid Century 
  ca_years(start = 2000, end = 2005)

# Check if request has no erros
locaschl_et_cap %>% ca_preflight()

head(locaschl_et_cap)
```

```{r}
#Next, we fetch the data. Because this is still a lot of locations, we’ll use ca_getvals_db() to save it in a database:
data_dir_precip_hist_85 <- tools::R_user_dir("pr_hist_85", which = "data")
schools_dir_precip_hist_85 <- file.path(data_dir_precip_hist_85, "schools_pr_cur_85") %>% normalizePath(mustWork = FALSE)
if (!file.exists(schools_dir_precip_hist_85)) dir.create(schools_dir_precip_hist_85, recursive = TRUE)

## Define a new SQLite file name
locaschl_fn_hist85 <- file.path(schools_dir_precip_hist_85, "loca_schl.sqlite") %>% normalizePath(mustWork = FALSE)


## Fetch data
locaschl_et_rtbl <- locaschl_et_cap %>%
  ca_getvals_db(db_fn = locaschl_fn_hist85, db_tbl = "locaschl_et", new_recs_only = TRUE, quiet = TRUE)


head(locaschl_et_rtbl)
```

```{r}
# Transform dataset 
locaschl_et_rtbl_1 <- locaschl_et_rtbl %>% 
  mutate(month = lubridate::month(dt),
         year = lubridate::year(dt),
         val = as.numeric(val))

# Set threshold determined in 'caladapt_precip_r.qmd' and calculate days over
locaschl_et_rtbl_1 <- locaschl_et_rtbl_1 %>%
   mutate(threshold = ifelse(val > 0.000216, 1, 0)) %>%
   # Find total number of days by month and year
   group_by(id, month, year) %>%
   summarise(count = sum(threshold))
# 
# # View results
 head(locaschl_et_rtbl_1)
```

```{r}
# # Find average monthly yearly days above threshold for 4 GCM's
 locaschl_et_rtbl_2 <- locaschl_et_rtbl_1 %>% 
   mutate(average = count/4)

# Rejoin with school points + LOCA id
 ca_schools_et_sf <- ca_schools_loca_sf %>%  
   left_join(locaschl_et_rtbl_2, by = "id", copy = TRUE) 

 head(ca_schools_et_sf)
```

```{r}
# Find yearly extreme precipitation
 ca_schools_et_year <- ca_schools_et_sf %>% 
   group_by(CDSCode, year) %>% 
   summarise(total = sum(count))%>% 
  st_drop_geometry()
#   
 # Find sum total of extreme precipitation days across 1,0008 schools from 2035-2064
 ca_schools <- ca_schools_et_year %>% 
   group_by(CDSCode) %>% 
   summarise(total_00_06 = sum(total)) %>% 
   mutate(total_00_06 = round(total_00_06,digits = 0))
```

```{r}
# write.csv(ca_schools, "/capstone/casaschools/precipitation/historic85.csv")
# 
# write.csv(ca_schools, "precipitation/historic85.csv")
# 
# #Export Extreme Precipitation Totals per year for Historic Scenario
# 
# write.csv(ca_schools_et_year, "/capstone/casaschools/precipitation/historic85_year.csv" )
# 
# write.csv(ca_schools_et_year, "precipitation/historic85_year.csv" )
```

# 
