---
title: "Accessing Extreme Heat Days for all Schools at RCP 8.5"
format: html
author: "CASAschools Team"
---

```{r}
# Libraries
library(dplyr)
library(tidyr)
library(sf)
library(lubridate)
library(caladaptr)
library(tidyverse)
```

Extreme heat day is calculated as maximum daily temperature days above 98F. This overall threshold was calculated in `extreme_heat/caladapt-r.qmd`. Determining the number of extreme heat days follows a similar approach. Still working with Cal-Adapt's Rpackage the first step is to aquire daily maximum temperature. Data is requested through Cal-Adapt's API request. Due to the size of the data and the instances of crashing the API request data was extracted at 30 year intervals for the different scenarios. This QMD is for accessing RCP8.5 data for 2006 to 2064.

```{r}
# Load school point data
school_points <- st_read("/capstone/casaschools/schools_data/California_Schools_2022-23/California_Schools_2022-23.shp")
```

```{r}
# Filter to schools are active and remove unnecessary columns
school_points <- school_points %>% filter(Status == "Active") %>% 
  select(c(CDSCode, DistrictNa, SchoolName))
```

## RCP 8.5

### OBSERVED / HISTORIC 2000-2005

```{r}

# Import LOCA grid 
locagrid_sf <- ca_locagrid_geom()

# Check CRS of both data
st_crs(school_points) == st_crs(locagrid_sf)

# Transform school points to LOCA CRS
school_points <- st_transform(school_points, crs = st_crs(locagrid_sf))

# Spatially join points and grid
ca_schools_loca_sf <- school_points %>% st_join(locagrid_sf) 
ca_schools_loca_sf %>% st_drop_geometry()


# Unique ID values
# we only have to query 1,619 school points
loca_ids_schools <- ca_schools_loca_sf %>% pull(id) %>% unique()


# Create a point layer for the 1,619 LOCA grid cells that contain schools
loca_ctr_sf <- locagrid_sf %>%
  filter(id %in% loca_ids_schools) %>%
  st_centroid()


# API request 
locaschl_et_cap <- ca_loc_sf(loc = loca_ctr_sf, idfld = "id") %>%
  ca_livneh(TRUE) %>%
  ca_cvar("tasmax") %>%
  ca_period("day") %>%
  # Mid Century 
  ca_years(start = 2000, end = 2005)

# Check if request has no erros
locaschl_et_cap %>% ca_preflight()


#Next, we fetch the data. Because this is still a lot of locations, we’ll use ca_getvals_db() to save it in a database:
heat_2000_2005 <- tools::R_user_dir("heat_2000s", which = "data")
heat_2000_dir <- file.path(heat_2000_2005, "schools") %>% normalizePath(mustWork = FALSE)
if (!file.exists(heat_2000_dir)) dir.create(heat_2000_dir, recursive = TRUE)

## Define a new SQLite file name
locaschl_fn <- file.path(heat_2000_dir, "loca_schl.sqlite") %>% normalizePath(mustWork = FALSE)


## Fetch data
locaschl_et_rtbl <- locaschl_et_cap %>%
  ca_getvals_db(db_fn = locaschl_fn, db_tbl = "locaschl_et", new_recs_only = TRUE, quiet = TRUE)

head(locaschl_et_rtbl_1)

# Transform dataset 
locaschl_et_rtbl_1 <- locaschl_et_rtbl %>% 
  mutate(month = lubridate::month(dt),
         year = lubridate::year(dt))

# Set threshold determined in 'caladapt-r.qmd' and calculate days over
locaschl_et_rtbl_1 <- locaschl_et_rtbl_1 %>%
  mutate(threshold = ifelse(val > 36.42294, 1, 0)) %>%
  # Find total number of days by month and year
  group_by(id, month, year) %>%
  summarise(count = sum(threshold))

head(locaschl_et_rtbl_1)


# Rejoin with school points + LOCA id
ca_schools_et_sf <- ca_schools_loca_sf %>% 
  left_join(locaschl_et_rtbl_1, by = "id", copy = TRUE) 

head(ca_schools_et_sf)


# Find yearly extreme heat days
ca_schools_et_year <- ca_schools_et_sf %>% 
  group_by(CDSCode, year) %>% 
  summarise(total = sum(count)) %>% 
  st_drop_geometry()
  
# Find sum total of extreme heat days across 1,0008 schools from 2035-2064
ca_schools <- ca_schools_et_year %>% 
  group_by(CDSCode) %>% 
  summarise(total_06_34 = sum(total)) %>% 
  mutate(total_06_34 = round(total_06_34,digits = 0))

head(ca_schools_et_year)

# Export file

# write.csv(ca_schools_et_year, "/capstone/casaschools/shiny_dashboard/data/extreme_heat/heat_2000_2005.csv")
```

### CURRENT CENTURY 2006 - 2034

```{r}

# Import LOCA grid 
locagrid_sf <- ca_locagrid_geom()

# Check CRS of both data
st_crs(school_points) == st_crs(locagrid_sf)

# Transform school points to LOCA CRS
school_points <- st_transform(school_points, crs = st_crs(locagrid_sf))

# Spatially join points and grid
ca_schools_loca_sf <- school_points %>% st_join(locagrid_sf) 
ca_schools_loca_sf %>% st_drop_geometry()


# Unique ID values
# we only have to query 1,619 school points
loca_ids_schools <- ca_schools_loca_sf %>% pull(id) %>% unique()


# Create a point layer for the 1,619 LOCA grid cells that contain schools
loca_ctr_sf <- locagrid_sf %>%
  filter(id %in% loca_ids_schools) %>%
  st_centroid()


# API request 
locaschl_et_cap <- ca_loc_sf(loc = loca_ctr_sf, idfld = "id") %>%
  # CanESM2, CNRM-CM5, HadGEM2-ES, MIROC5
  ca_gcm(gcms[1:4]) %>%
  ca_scenario(c("rcp85")) %>%
  ca_cvar("tasmax") %>%
  ca_period("day") %>%
  # Mid Century 
  ca_years(start = 2006, end = 2034)

# Check if request has no erros
locaschl_et_cap %>% ca_preflight()


#Next, we fetch the data. Because this is still a lot of locations, we’ll use ca_getvals_db() to save it in a database:
heat_2006_20035 <- tools::R_user_dir("heat_rcp85_2006_2034", which = "data")
schools_heat_rcp85_2006 <- file.path(heat_2006_20035, "schools") %>% normalizePath(mustWork = FALSE)
if (!file.exists(schools_heat_rcp85_2006)) dir.create(schools_heat_rcp85_2006, recursive = TRUE)

## Define a new SQLite file name
locaschl_fn <- file.path(schools_heat_rcp85_2006, "loca_schl.sqlite") %>% normalizePath(mustWork = FALSE)


## Fetch data
locaschl_et_rtbl <- locaschl_et_cap %>%
  ca_getvals_db(db_fn = locaschl_fn, db_tbl = "locaschl_et", new_recs_only = TRUE, quiet = TRUE)

# Transform dataset 
locaschl_et_rtbl_1 <- locaschl_et_rtbl %>% 
  mutate(month = lubridate::month(dt),
         year = lubridate::year(dt))

# Set threshold determined in 'caladapt-r.qmd' and calculate days over
locaschl_et_rtbl_1 <- locaschl_et_rtbl_1 %>%
  mutate(threshold = ifelse(val > 309.817, 1, 0)) %>%
  # Find total number of days by month and year
  group_by(id, month, year) %>%
  summarise(count = sum(threshold))

# Find average monthly yearly days above threshold for 4 GCM's
locaschl_et_rtbl_2 <- locaschl_et_rtbl_1 %>% 
  mutate(avergae = count/4)

# Rejoin with school points + LOCA id
ca_schools_et_sf <- ca_schools_loca_sf %>% 
  left_join(locaschl_et_rtbl_2, by = "id", copy = TRUE) 

head(ca_schools_et_sf)


# Find yearly extreme heat days
ca_schools_et_year <- ca_schools_et_sf %>% 
  group_by(CDSCode, year) %>% 
  summarise(total = sum(avergae))
  
# Find sum total of extreme heat days across 1,0008 schools from 2035-2064
ca_schools <- ca_schools_et_year %>% 
  group_by(CDSCode) %>% 
  summarise(total_06_34 = sum(total)) %>% 
  mutate(total_06_34 = round(total_06_34,digits = 0))


# Export file

# write.csv(ca_schools_et_year, "/capstone/casaschools/extreme_heat/rcp8.5_2006_2034.csv")
```

### MID CENTURY 2035 - 2064

```{r}
# Import LOCA grid 
locagrid_sf <- ca_locagrid_geom()

# Check CRS of both data
st_crs(school_points) == st_crs(locagrid_sf)

# Transform school points to LOCA CRS
school_points <- st_transform(school_points, crs = st_crs(locagrid_sf))

# Spatially join points and grid
ca_schools_loca_sf <- school_points %>% st_join(locagrid_sf) 
ca_schools_loca_sf %>% st_drop_geometry()


# Unique ID values
# we only have to query 1,619 school points
loca_ids_schools <- ca_schools_loca_sf %>% pull(id) %>% unique()


# Create a point layer for the 1,619 LOCA grid cells that contain schools
loca_ctr_sf <- locagrid_sf %>%
  filter(id %in% loca_ids_schools) %>%
  st_centroid()


# API request 
locaschl_et_cap <- ca_loc_sf(loc = loca_ctr_sf, idfld = "id") %>%
  # CanESM2, CNRM-CM5, HadGEM2-ES, MIROC5
  ca_gcm(gcms[1:4]) %>%
  ca_scenario(c("rcp85")) %>%
  ca_cvar("tasmax") %>%
  ca_period("day") %>%
  # Mid Century 
  ca_years(start = 2035, end = 2064)

# Check if request has no erros
locaschl_et_cap %>% ca_preflight()


#Next, we fetch the data. Because this is still a lot of locations, we’ll use ca_getvals_db() to save it in a database:
heat_rcp85_35_64 <- tools::R_user_dir("heat_rcp85_2035_2064", which = "data")
heat_rcp85_dir <- file.path(heat_rcp85_35_64, "schools") %>% normalizePath(mustWork = FALSE)
if (!file.exists(heat_rcp85_dir)) dir.create(heat_rcp85_dir, recursive = TRUE)

## Define a new SQLite file name
locaschl_fn <- file.path(heat_rcp85_dir, "loca_schl.sqlite") %>% normalizePath(mustWork = FALSE)


## Fetch data
locaschl_et_rtbl <- locaschl_et_cap %>%
  ca_getvals_db(db_fn = locaschl_fn, db_tbl = "locaschl_et", new_recs_only = TRUE, quiet = TRUE)

# Transform dataset 
locaschl_et_rtbl_1 <- locaschl_et_rtbl %>% 
  mutate(month = lubridate::month(dt),
         year = lubridate::year(dt),
         val = as.integer(val))

# Set threshold determined in 'caladapt-r.qmd' and calculate days over
locaschl_et_rtbl_1 <- locaschl_et_rtbl_1 %>%
  mutate(threshold = ifelse(val > 309.817, 1, 0)) %>%
  # Find total number of days by month and year
  group_by(id, month, year) %>%
  summarise(count = sum(threshold))

# View results
head(locaschl_et_rtbl_1)

# Find average monthly yearly days above threshold for 4 GCM's
locaschl_et_rtbl_2 <- locaschl_et_rtbl_1 %>% 
  mutate(avergae = count/4)

# Rejoin with school points + LOCA id
ca_schools_et_sf <- ca_schools_loca_sf %>% 
  left_join(locaschl_et_rtbl_2, by = "id", copy = TRUE) 

head(ca_schools_et_sf)



# Find yearly extreme heat days
ca_schools_et_year <- ca_schools_et_sf %>% 
  group_by(CDSCode, year) %>% 
  summarise(total = sum(avergae)) %>% 
  st_drop_geometry()
  
# Find sum total of extreme heat days across 1,0008 schools from 2035-2064
ca_schools <- ca_schools_et_year %>% 
  group_by(CDSCode) %>% 
  summarise(total_35_64 = sum(total)) %>% 
  mutate(total_35_64 = round(total_35_64,digits = 0))


# Export file
# write.csv(ca_schools, "/capstone/casaschools/extreme_heat/rcp8.5_2006_2034.csv")
# 
# 
# write.csv(ca_schools, "/Users/hazelvaquero/CASAschools/climate_hazards/extreme_heat/output_data/ca_schools_rcp8.5_2035_2064.csv")


```

### Combine datasets

```{r}
options(scipen = 999)


heat_2000_2005 <- read.csv("/capstone/casaschools/extreme_heat/heat_2000_2005.csv")


rcp85_2006_2034 <- read.csv("/capstone/casaschools/extreme_heat/rcp8.5_2006_2034.csv")

rcp85_2035_2064 <- read.csv("/capstone/casaschools/extreme_heat/rcp8.5_2035_2064.csv")


## Combine to make rcp45 2000-2064
heat_rcp85_2000_2064 <- rbind(heat_2000_2005, 
                              rcp85_2006_2034, 
                              rcp85_2035_2064)

heat_rcp85_2000_2064 <- heat_rcp85_2000_2064 %>%
  select(-X)

### export csv to shinydashboard data
# write.csv(heat_rcp85_2000_2064, "/capstone/casaschools/shiny_dashboard/data/extreme_heat/heat_rcp85_2000_20064.csv")
# 
# write.csv(heat_rcp85_2000_2064_zeros, "/capstone/casaschools/shiny_dashboard/data/extreme_heat/heat_rcp85_2000_20064_zeros.csv")
# 
# write.csv(heat_rcp85_2000_2064, "/capstone/casaschools/extreme_heat/heat_rcp85_2000_20064.csv")
```

### Data preparation for dashboard

```{r}
 options(scipen = 999)

# Combine rcp4.5 & rcp8.5
# Load RCP 4.5
rcp45 <- read_csv("/capstone/casaschools/extreme_heat/heat_rcp45_2000_20064.csv")

# Load RCP 8.5
rcp85 <- read_csv("/capstone/casaschools/extreme_heat/heat_rcp85_2000_20064.csv")

# Add extra columns to distinguish them
rcp45 <- rcp45 %>% 
  mutate(scenario = "Reduced greenhouse gas emissions",
         total = round(total)) %>% 
  select(c(CDSCode, year, total, scenario))


rcp85 <- rcp85 %>% 
  mutate(scenario = "High greenhouse gas emissions",
         total = round(total)) %>% 
  select(c(CDSCode, year, total, scenario))

# Join datasets
extreme_heat <- bind_rows(rcp45, rcp85)

extreme_heat <- extreme_heat %>% 
  # Transform CDSCode as character
  mutate(CDSCode_old = as.character(CDSCode),
         # Count how many characters
         count = nchar(CDSCode),
         # If 13 add a 0 in front
         CDSCode = ifelse(count == 13, paste0("0",CDSCode_old),
                          CDSCode_old)) %>% 
  select(-c(CDSCode_old, count)) %>% 
  filter(year > 2005)

```

Each school has a unique 14 digit associated CDSCode. When working with the California_Schools_2022-23 dataset the `CDSCode` loses its initial `0`. To prepare the extreme heat data to be joined with the rest of the hazards this missing `0` needs to be added back.

```{r}
options(scipen = 999)

# remove geometry of school points
school_points <- school_points %>% st_drop_geometry()

# Merge extreme heat with school names
extreme_heat <- left_join(extreme_heat, school_points, by = "CDSCode") 

# Check if 10,008 schools are there
length(unique(extreme_heat$CDSCode))

# Export csv
# write.csv(extreme_heat, "/capstone/casaschools/shiny_dashboard/data/extreme_heat/extreme_heat.csv")
# 
# 
# ex_h <- read.csv("/capstone/casaschools/shiny_dashboard/data/extreme_heat/extreme_heat.csv")
```
