```{r, message = FALSE}
library(sf)
library(tidyverse)
library(terra)
library(leaflet)
```

Mapping wildfire for school districts. The plan to map is to overlay school buffers areas on top of the wildfire raster data. We should use the buffers that are the complete buffers for mapping so that users can see information about the school's average wildfire hazard potential score.


```{r}
# read in ca wildfire data
whp_ca <- rast("/capstone/casaschools/wildfire/intermediate_layers/whp_ca.tif")

# read in reclassified whp data
whp_reclass <- rast("/capstone/casaschools/wildfire/intermediate_layers/whp_reclass.tif")

# read in archived reclassified whp data
whp_reclass2012 <- rast("/capstone/casaschools/wildfire/intermediate_layers/whp_reclass2012.tif")

# read in school buffers clipped
schools_buffers_clip <- st_read("/capstone/casaschools/schools_data/schools_buffer_clip/schools_buffer_clip.shp",
                                quiet = TRUE)

# read in full school buffers
schools_buffers <- st_read("/capstone/casaschools/schools_data/schools_buffer/schools_points_buffer.shp",
                           quiet = TRUE)
```


```{r}
## mapping the 2023 whp data
# filter for one school
dp <- schools_buffers %>% 
  filter(SchoolName == "Dos Pueblos Senior High")

# create the school point as the centroid of the buffer
dp_point <- dp %>% 
  st_centroid()

# select the wildfire hazard potential cells that overlap
whp_school <- crop(whp_reclass, dp)
whp_school2012 <- crop(whp_reclass2012, dp)

# data needs to be in lot/long format, or transformed to WGS 1984
dp <- st_transform(dp, crs = 4326)
dp_point <- st_transform(dp_point, crs = 4326)

# define color palette and labels for wildfire hazard potential
labels <- c("developed or open water", "very low", "low", "moderate", "high", "very high", "")
whp_colors <- c("grey", "#fee391", "#fec44f", "#fe9929", "#d95f0e", "#993404", "transparent")
whp_palette <- colorFactor(palette = whp_colors,
                           domain = values(whp_school),
                           na.color = "transparent")

# leaflet map of wildfire hazard potential for 2023
leaflet() %>% 
  
  # add topo basemap
  addProviderTiles(providers$Esri.WorldTopoMap, group = "topographic map") %>%
  
  # add imagery basemap
  addProviderTiles(providers$Esri.WorldImagery, group = "satellite imagery") %>%
  
  # add cropped wildfire hazard potential raster for 2023 
  addRasterImage(whp_school, colors = whp_palette, opacity = .7, group = "wildfire hazard potential") %>%
  
  # add school buffer polygon
  addPolygons(data = dp, color = "black", fill = FALSE, 
              weight = 2, group = "school community area") %>% 
  
  # add school point
  addCircleMarkers(data = dp_point, color = "blue", stroke = FALSE, 
                   weight = 3, radius = 5, fillOpacity = 1,
                   group = "school point") %>% 
  
  # add legend for wildfire hazard potential with custom labels
  addLegend("bottomright", colors = rev(whp_colors), labels = rev(labels),
            title = "wildfire hazard potential", opacity = 1) %>% 
  
  # add option to toggle data on and off
  addLayersControl(
    overlayGroups = c("wildfire hazard potential", "school community area", "school point"),
    baseGroups = c("topographic map", "satellite imagery"),
    options = layersControlOptions(collapsed = FALSE))
```

```{r}
# leaflet map of wildfire hazard potential for 2012
leaflet() %>% 
  
   # add topo basemap
  addProviderTiles(providers$Esri.WorldTopoMap, group = "topographic map") %>%
  
  # add imagery basemap
  addProviderTiles(providers$Esri.WorldImagery, group = "satellite imagery") %>% 
  
  # add cropped wildfire hazard potential raster for 2023 
  addRasterImage(whp_school2012, colors = whp_palette, opacity = .7, group = "wildfire hazard potential") %>%
  
  # add school buffer polygon
  addPolygons(data = dp, color = "black", fill = FALSE, 
              weight = 2, group = "school community area") %>% 
  
  # add school point
  addCircleMarkers(data = dp_point, color = "blue", stroke = FALSE, 
                   weight = 3, radius = 5, fillOpacity = 1,
                   group = "school point") %>% 
  
  # add legend for wildfire hazard potential with custom labels
   addLegend("bottomright", colors = rev(whp_colors), labels = rev(labels),
            title = "wildfire hazard potential", opacity = 1) %>% 
  
   # add option to toggle data on and off
  addLayersControl(
    overlayGroups = c("wildfire hazard potential", "school community area", "school point"),
    baseGroups = c("topographic map", "satellite imagery"),
    options = layersControlOptions(collapsed = FALSE))
```

### testing

removing NA values entirely so they are just holes in the data
```{r}
mask <- !is.na(whp_school2012)

test_mask <- mask(whp_school2012, mask)

# define color palette and labels for wildfire hazard potential
labels2 <- c("developed or open water", "very low", "low", "moderate", "high", "very high")
whp_colors2 <- c("grey", "#fee391", "#fec44f", "#fe9929", "#d95f0e", "#993404")
whp_palette2 <- colorFactor(palette = whp_colors,
                           domain = values(whp_school))

# leaflet map of wildfire hazard potential for 2012
leaflet() %>% 
  
   # add topo basemap
  addProviderTiles(providers$Esri.WorldTopoMap, group = "topographic map") %>%
  
  # add imagery basemap
  addProviderTiles(providers$Esri.WorldImagery, group = "satellite imagery") %>% 
  
  # add cropped wildfire hazard potential raster for 2023 
  addRasterImage(test_mask, colors = whp_palette2, opacity = .7, group = "wildfire hazard potential") %>%
  
  # add school buffer polygon
  addPolygons(data = dp, color = "black", fill = FALSE, 
              weight = 2, group = "school community area") %>% 
  
  # add school point
  addCircleMarkers(data = dp_point, color = "blue", stroke = FALSE, 
                   weight = 3, radius = 5, fillOpacity = 1,
                   group = "school point") %>% 
  
  # add legend for wildfire hazard potential with custom labels
   addLegend("bottomright", colors = rev(whp_colors2), labels = rev(labels2),
            title = "wildfire hazard potential", opacity = 1) %>% 
  
   # add option to toggle data on and off
  addLayersControl(
    overlayGroups = c("wildfire hazard potential", "school community area", "school point"),
    baseGroups = c("topographic map", "satellite imagery"),
    options = layersControlOptions(collapsed = FALSE))
```

57% of cells in the 2012 and 2023 reclassified raster are NA. 52% of cells in the original 2012 data are NA. 43% of cells in the original us data are NA  
```{r}
# 2012 data
((global(whp_us2012, fun = "isNA")[1, 1]) / ncell(whp_us2012)) * 100

(global(whp_reclass2012, fun = "isNA") / ncell(whp_reclass2012)) * 100

# for 2023 data
(global(whp_us, fun = "isNA") / ncell(whp_us)) * 100

(global(whp_reclass, fun = "isNA") / ncell(whp_reclass)) * 100
```

