```{r, message = FALSE}
library(sf)
library(tidyverse)
library(tmap)
library(terra)
```

```{r}
# read in wildfire data
whp_us <- rast("/capstone/casaschools/wildfire/raw_data/whp2023_cls_conus.tif")

whp_ca <- rast("/capstone/casaschools/wildfire/intermediate_layers/whp_ca.tif")

# read in school buffers
ca_schools <- st_read("/capstone/casaschools/schools_data/schools_buffer_clip/schools_buffer_clip.shp",
                      quiet = TRUE)

# read in California boundary
ca_boundary <- st_read("/capstone/casaschools/schools_data/CA_boundary/CA_boundary.shp",
                       quiet = TRUE)
```


## clip wildfire raster to California

The wildfire hazard potential layer is mapped for the entire United States. For the purposes of our analysis, we only need the area that covers California. We first project the wildfire hazard potential raster to be in the same CRS as the California boundary. We then crop the raster to the boundary with the argument mask = TRUE. This crops the raster to the exact boundary, not just the bounding box of the California layer. This cropping causes fractional values to occur in our raster, which we don't want. This arises from cells on the edges not aligning exactly with the border being resampled. So, we round all the values in the raster to get back on a categorical scale from 1-7. Then, we write the file to the server

```{r}
# convert ca_boundary sf object to a terra vector
#ca_vect <- vect(ca_boundary)

# reproject the whp raster to be on the same coordinate reference system as the California polygon
whp_us <- project(whp_us, ca_boundary)

summary(whp_us)

# check that the crs was changed
#st_crs(whp_us) == st_crs(ca_schools)

# crop to the extent of the California boundary
#whp_ca <- crop(whp_us, ca_boundary, mask = TRUE, touches = FALSE)

whp_ca <- crop(whp_us, ca_boundary, mask = TRUE, touches = TRUE)

whp_ca <- round(whp_ca)

#unique(whp_ca)

# save output shapefile as an intermediate layer
writeRaster(whp_ca, "/capstone/casaschools/wildfire/intermediate_layers/whp_ca.tif")
```


## assign wildfire hazard potential values to school buffers

From the metadata, the classifications in the data are as follows:

Values:
1- very low
2- low
3- moderate
4- high
5- very high
6- non-burnable
7- water

Our process for assigning values to school buffers will be to take the average value of cells that overlap with each school buffer. The categories "non-burnable" include developed areas not at risk of wildfire, and "water" also isn't at risk of burning. So, we'll reclassify these values as 0. This essentially says that they are not burnable at all, and represent no risk to the area. The average value of cells will be rounded to the nearest whole number, giving each school area a final value between 1-5, representing very low, low, moderate, high, and very high risk categories.

- How to treat NA values?

```{r}
## reclassify the whp raster values

# remove NA values from the raster

# create a reclassification matrix
rcl_table <- matrix(c(1, 1,
                      2, 2,
                      3, 3,
                      4, 4,
                      5, 5,
                      6, 0,
                      7, 0), ncol=2, byrow=TRUE)

# reclassify the whp raster
whp_reclass <- classify(whp_ca, rcl = rcl_table)

# check the values of the raster
summary(whp_reclass)
```

### calculating the value of wildfire hazard potential for every school buffer

The next step in this process is to assign wildfire hazard potential scores for every buffer. This involves taking every cell that overlaps with the school buffers and averaging the wildfire hazard potential score. We'll include cells on the boundary without interpolating scores based on percentage of area. The rationale for this is that score corresponds to the entire cell. If a cell has a wildfire hazard potential score of 2, a buffer boundary cutting it in half doesn't also divide the score by 2. The entire underlying area has a score of 2.

```{r}
# test with a subset first
# select one school
envision <- ca_schools %>% 
  filter(FedID == "060161410947")

# select overlapping raster cells
#whp_envision <- crop(whp_reclass, envision)

whp_envision2 <- crop(whp_reclass, envision, mask = TRUE)

plot(whp_envision2)

## convert to polygons to manually calculate mean
whp_test_poly <- as.polygons(whp_envision2, values = TRUE, aggregate = FALSE) %>% 
  st_as_sf()

envision_mean_test <- mean(whp_test_poly$Band_1)

# plot to see that it worked
tm_shape(envision) +
  tm_polygons(fill = NA) +
  tm_shape(whp_envision2) +
  tm_raster()

# extract the mean value of all cells that intersect with the buffer
# touches = TRUE grabs all of the cells that intersect, not just those fully contained within the polygon
# bind = TRUE attaches the polygon information to the calculated mean
envision_mean <- extract(whp_envision2, envision, fun = "mean", method = "simple", ID = TRUE, touches = TRUE, bind = TRUE) %>% 
  st_as_sf()
```


#### doing it for the whole dataset

```{r}
# crop the whp raster to each school, using the buffers as a mask
whp_schools_mask <- crop(whp_reclass, ca_schools, mask = TRUE)

# plot the result to check
# tm_shape(whp_schools) +
#   tm_raster()

# calculate the mean whp value of every school buffer
# our function is mean, we remove NA values
# touches = TRUE ensures that any cells that intersect, instead of being fully contained, are included
# bind = TRUE binds the result to the 
start <- proc.time()

whp_schools <- extract(whp_schools_mask, ca_schools, fun = "mean", na.rm = TRUE, 
                       method = "simple", touches = TRUE, bind = TRUE) %>% 
  st_as_sf()

end <- proc.time() - start

# view the min and the max to see that it's between 0-5
min(whp_schools$Band_1)
max(whp_schools$Band_1)

# find the number of rows that are NA
sum(is.nan(whp_schools$Band_1))

# round the values to the nearest whole number
whp_schools <- whp_schools %>% 
  rename(whp = Band_1) %>% 
  mutate(whp = round(whp, digits = 0),
         whp = ifelse(whp == 0, 1, whp))

# view the min and max
min(whp_schools$whp)
max(whp_schools$whp)
```


### testing 

```{r}
tmap_mode("view")

tm_shape(whp_ca) +
  tm_raster()

tm_shape(whp_reclass) +
  tm_raster()
```

### testing assumptions

```{r}
## find the number of cells that become fractional after cropping to California
# total number of cells
# this tells us 
whp_test <- whp_ca_round - whp_ca

# find how many might have been changed to a different score
selected_cells <- (whp_test < -0.499) | (whp_test > 0.499)

plot(selected_cells)

true_count = sum(values(selected_cells), na.rm = TRUE)

## find the number of cells that changed value that overlap with school buffers
```

